Effects of Antenna Beam Chromaticity on Redshifted 21~cm Power Spectrum and Implications for Hydrogen Epoch of Reionization Array

### 2016-03-15 ForegroundSimulations/ms.tex
Replace use of "sensitivity" with "dynamic range"
From 10a4726 (fgsim):
====================

In ms.tex,

(*) Replace usage of "sensitivity" with "dynamic range" throughout

This paper is organized as follows. \S\ref{sec:HERA} introduces the HERA instrument. A brief summary of the delay spectrum technique used extensively in this analysis and the recently confirmed {\it pitchfork} effect are presented in \S\ref{sec:delay-spectrum}. \S\ref{sec:sim} describes foreground simulations including antenna beam pattern, all-sky foreground models and two independent EoR models. \S\ref{sec:beam-chromaticity} investigates the effects of chromaticity of antenna beam on the resulting delay power spectrum and the cosmologically motivated constraints it places on reflections in the instrument. Foreground-limited sensitivity of HERA under simple {\it foreground avoidance} criteria are demonstrated in \S\ref{sec:eor-sensitivity}. Our findings are summarized in \S\ref{sec:summary}.

@@

\section{Foreground-limited EoR Sensitivity}\label{sec:eor-sensitivity}

Here our goal is to estimate the foreground-limited sensitivity (ratio of EoR signal power to foreground power) in the presence of antenna beam chromaticity, using HERA as an example. As stated earlier, we estimate this ratio for HERA in the most conservative scenario where we consider only {\it foreground avoidance} under simple spectral weighting scheme. We emphasize that other strategies involving {\it foreground removal} and inverse covariance-based weighting schemes will produce more optimal results than presented here.

In order to avoid signal evolution across the entire bandwidth, we use subbands in which EoR signal evolution is not expected to be significant. These subbands with $B_\textrm{eff}=10$~MHz are defined by spectral weights obtained by convolution of {\it Blackman-Harris} window with itself. These subbands are centered at 150~MHz and 170~MHz to estimate EoR-foreground power ratio at $z \approx 8.47$ and $z \approx 7.36$ respectively. We use both EoR models 1 and 2 to reduce the dependence of this analysis on any single EoR model.

@@

\begin{figure*}[htb]
\centering
\subfloat[][150~MHz subband ($z \approx 8.47$)]{\label{fig:asm-dps-beam-chromaticity-baselines-150-subband}\includegraphics[width=\linewidth]{sim_asm_foreground_eor_beam_chromaticity_150.0_MHz_subband_win_method_bhw2.0.eps}} \\
\subfloat[][170~MHz subband ($z \approx 7.36$)]{\label{fig:asm-dps-beam-chromaticity-baselines-170-subband}\includegraphics[width=\linewidth]{sim_asm_foreground_eor_beam_chromaticity_170.0_MHz_subband_win_method_bhw2.0.eps}}
\caption{EoR signal and foreground delay power spectrum in units of K$^2 (h^{-1}$~Mpc$)^3$ in 150~MHz (top) and 170~MHz (bottom) subbands ($B_\textrm{eff}=10$~MHz) on eight unique baseline lengths of HERA-19 at arbitrary sky pointings. The length and orientation of the baseline vector corresponding to each panel is annotated on the top right corner. EoR models 1 and 2 are shown in cyan and gray respectively. The foregrounds obtained with achromatic, {\it Airy} and simulated chromatic antenna beams are shown in black, red and blue respectively. EoR-foreground power ratio is highest for antenna beam with least chromaticity and vice versa. Even for the simulated chromatic antenna beam pattern, which has the highest chromaticity among the antenna beam models considered, foreground power will be lower than signal power from the two independent EoR models by more than two orders of magnitude for $|k_\parallel| \gtrsim 0.2\,h$~Mpc$^{-1}$ on all HERA baselines.}
\label{fig:subbands}
\end{figure*}

@@

We investigate the EoR-foreground power ratio in two dimensional parameter space formed by baselines and pointings in Right Ascension (RA$\equiv\alpha$) to highlight capabilities of HERA and provide clues for the best observing window.

@@

We estimate the worst case EoR-foreground power ratio in the $\boldsymbol{b}$--$\alpha$ parameter space as:
\begin{align}
  \rho(\boldsymbol{b},f_0,\alpha) &= \min_{|k_\parallel|>k_\parallel^\textrm{FA}(|\boldsymbol{b}|)}\left\{\frac{P_\textrm{H{\sc i}}(|\boldsymbol{b}|,k_\parallel,f_0)}{P_\textrm{FG}(\boldsymbol{b},k_\parallel,f_0,\alpha)}\right\}.
\end{align}
This denotes the minimum EoR-foreground power ratio in $k_\parallel$-modes of interest for {\it foreground avoidance}, $|k_\parallel| > k_\parallel^\textrm{FA}(\boldsymbol{b})$, for each baseline vector ($\boldsymbol{b}$), subband ($f_0$), and sky pointing ($\alpha$).

@@

\begin{figure*}[htb]
\centering
\subfloat[][EoR model 1, 150~MHz subband ]{\label{fig:21cmfast-fg-ratio-150}\includegraphics[width=0.45\linewidth]{eor_fg_ratio_llim_21cmfast_chrmbeam_sim_150.0_MHz_subband.eps}}
\subfloat[][EoR model 2, 150~MHz subband ]{\label{fig:lidz-fg-ratio-150}\includegraphics[width=0.45\linewidth]{eor_fg_ratio_llim_lidz_chrmbeam_sim_150.0_MHz_subband.eps}} \\
\subfloat[][EoR model 1, 170~MHz subband ]{\label{fig:21cmfast-fg-ratio-170}\includegraphics[width=0.45\linewidth]{eor_fg_ratio_llim_21cmfast_chrmbeam_sim_170.0_MHz_subband.eps}}
\subfloat[][EoR model 2, 170~MHz subband ]{\label{fig:lidz-fg-ratio-170}\includegraphics[width=0.45\linewidth]{eor_fg_ratio_llim_lidz_chrmbeam_sim_170.0_MHz_subband.eps}}
\caption{Worst case EoR-foreground power ratio, $\rho(\boldsymbol{b},f_0,\alpha)$, in {\it foreground avoidance} modes for different HERA-19 baselines as a function of RA ($\alpha$) of pointing on the sky. Each panel corresponds to the EoR model and subband indicated. The baseline vector is specified by its length (in m) and orientation (in degrees) at the bottom and top of the $x$-axis respectively. The color scale is logarithmic. Even in the worst case, HERA will be able detect EoR drawn from independent models on all baselines for a large fraction of a full 24~hour sky transit cycle. $\rho(\boldsymbol{b},f_0,\alpha)$ is seen to increase with increasing baseline length. This is because the loss in sensitivity to diffuse foreground power with increase in antenna spacings is faster than in the case of the EoR signal.}
\label{fig:eor-fg-ratios}
\end{figure*}

>>>>

This paper is organized as follows. \S\ref{sec:HERA} introduces the HERA instrument. A brief summary of the delay spectrum technique used extensively in this analysis and the recently confirmed {\it pitchfork} effect are presented in \S\ref{sec:delay-spectrum}. \S\ref{sec:sim} describes foreground simulations including antenna beam pattern, all-sky foreground models and two independent EoR models. \S\ref{sec:beam-chromaticity} investigates the effects of chromaticity of antenna beam on the resulting delay power spectrum and the cosmologically motivated constraints it places on reflections in the instrument. EoR-foreground dynamic range of HERA under simple {\it foreground avoidance} criteria are demonstrated in \S\ref{sec:eor-sensitivity}. Our findings are summarized in \S\ref{sec:summary}.

@@

\section{EoR-Foreground Dynamic Range}\label{sec:eor-sensitivity}

Here our goal is to estimate the EoR-foreground dynamic range (ratio of EoR signal power to foreground power) in the presence of antenna beam chromaticity, using HERA as an example. As stated earlier, we estimate this ratio for HERA in the most conservative scenario where we consider only {\it foreground avoidance} under simple spectral weighting scheme. We emphasize that other strategies involving {\it foreground removal} and inverse covariance-based weighting schemes will produce more optimal results than presented here.

In order to avoid signal evolution across the entire bandwidth, we use subbands in which EoR signal evolution is not expected to be significant. These subbands with $B_\textrm{eff}=10$~MHz are defined by spectral weights obtained by convolution of {\it Blackman-Harris} window with itself. These subbands are centered at 150~MHz and 170~MHz to estimate EoR-foreground dynamic range at $z \approx 8.47$ and $z \approx 7.36$ respectively. We use both EoR models 1 and 2 to reduce the dependence of this analysis on any single EoR model.

@@

\begin{figure*}[htb]
\centering
\subfloat[][150~MHz subband ($z \approx 8.47$)]{\label{fig:asm-dps-beam-chromaticity-baselines-150-subband}\includegraphics[width=\linewidth]{sim_asm_foreground_eor_beam_chromaticity_150.0_MHz_subband_win_method_bhw2.0.eps}} \\
\subfloat[][170~MHz subband ($z \approx 7.36$)]{\label{fig:asm-dps-beam-chromaticity-baselines-170-subband}\includegraphics[width=\linewidth]{sim_asm_foreground_eor_beam_chromaticity_170.0_MHz_subband_win_method_bhw2.0.eps}}
\caption{EoR signal and foreground delay power spectrum in units of K$^2 (h^{-1}$~Mpc$)^3$ in 150~MHz (top) and 170~MHz (bottom) subbands ($B_\textrm{eff}=10$~MHz) on eight unique baseline lengths of HERA-19 at arbitrary sky pointings. The length and orientation of the baseline vector corresponding to each panel is annotated on the top right corner. EoR models 1 and 2 are shown in cyan and gray respectively. The foregrounds obtained with achromatic, {\it Airy} and simulated chromatic antenna beams are shown in black, red and blue respectively. EoR-foreground dynamic range is highest for antenna beam with least chromaticity and vice versa. Even for the simulated chromatic antenna beam pattern, which has the highest chromaticity among the antenna beam models considered, foreground power will be lower than signal power from the two independent EoR models by more than two orders of magnitude for $|k_\parallel| \gtrsim 0.2\,h$~Mpc$^{-1}$ on all HERA baselines.}
\label{fig:subbands}
\end{figure*}

@@

We investigate the EoR-foreground dynamic range in two dimensional parameter space formed by baselines and pointings in Right Ascension (RA$\equiv\alpha$) to highlight capabilities of HERA and provide clues for the best observing window.

@@

We estimate the worst case EoR-foreground dynamic range in the $\boldsymbol{b}$--$\alpha$ parameter space as:
\begin{align}
  \rho(\boldsymbol{b},f_0,\alpha) &= \min_{|k_\parallel|>k_\parallel^\textrm{FA}(|\boldsymbol{b}|)}\left\{\frac{P_\textrm{H{\sc i}}(|\boldsymbol{b}|,k_\parallel,f_0)}{P_\textrm{FG}(\boldsymbol{b},k_\parallel,f_0,\alpha)}\right\}.
\end{align}
This denotes the minimum EoR-foreground dynamic range in $k_\parallel$-modes of interest for {\it foreground avoidance}, $|k_\parallel| > k_\parallel^\textrm{FA}(\boldsymbol{b})$, for each baseline vector ($\boldsymbol{b}$), subband ($f_0$), and sky pointing ($\alpha$).

@@

\begin{figure*}[htb]
\centering
\subfloat[][EoR model 1, 150~MHz subband ]{\label{fig:21cmfast-fg-ratio-150}\includegraphics[width=0.45\linewidth]{eor_fg_ratio_llim_21cmfast_chrmbeam_sim_150.0_MHz_subband.eps}}
\subfloat[][EoR model 2, 150~MHz subband ]{\label{fig:lidz-fg-ratio-150}\includegraphics[width=0.45\linewidth]{eor_fg_ratio_llim_lidz_chrmbeam_sim_150.0_MHz_subband.eps}} \\
\subfloat[][EoR model 1, 170~MHz subband ]{\label{fig:21cmfast-fg-ratio-170}\includegraphics[width=0.45\linewidth]{eor_fg_ratio_llim_21cmfast_chrmbeam_sim_170.0_MHz_subband.eps}}
\subfloat[][EoR model 2, 170~MHz subband ]{\label{fig:lidz-fg-ratio-170}\includegraphics[width=0.45\linewidth]{eor_fg_ratio_llim_lidz_chrmbeam_sim_170.0_MHz_subband.eps}}
\caption{Worst case EoR-foreground dynamic range, $\rho(\boldsymbol{b},f_0,\alpha)$, in {\it foreground avoidance} modes for different HERA-19 baselines as a function of RA ($\alpha$) of pointing on the sky. Each panel corresponds to the EoR model and subband indicated. The baseline vector is specified by its length (in m) and orientation (in degrees) at the bottom and top of the $x$-axis respectively. The color scale is logarithmic. Even in the worst case, HERA will be able detect EoR drawn from independent models on all baselines for a large fraction of a full 24~hour sky transit cycle. $\rho(\boldsymbol{b},f_0,\alpha)$ is seen to increase with increasing baseline length. This is because the loss in sensitivity to diffuse foreground power with increase in antenna spacings is faster than in the case of the EoR signal.}
\label{fig:eor-fg-ratios}
\end{figure*}
